### 2016-02-04 inputs/discussion.tex
phrasing

\subsection{Missing rate and the ``curse of dimensionality''}
\label{sec:missing_rate_curse_dimensionality}
In general, imputation methods were relatively robust to increasing missing rates. Only KNNI, and to a lesser extent RFI, showed slowly degrading performances at very high missing rates. The divergent response of the different imputation methods to missing rate became apparent only in the most challenging scenarios, when markers with up to 40-70\% missing genotypes were allowed in the data set.\\
KNNI's decrease in imputation accuracy with increasing missing rates can be interpreted in relation to the phenomenon known as the ``curse of dimensionality'' \cite{bellman_dynamic_1957,marimont_nearest_1979}: with the same amount of data, the increasing number of parameters - the ``dimensions'' of the problem - increases and complicates the identification of k neighbors which are close enough to the data point to be classified/imputed. 
On average, the side $l$ of the hypercube to include $k$ neighbors is a function of $k$, $n$ (sample size) and $p$ (number of parameters): $l = \left ( \frac{k}{n} \right )^{1/p}$. With $n$ constant, the hypercube in which the $k$ neighbors lie gets bigger as $p$ increases. This holds especially for local learning methods (e.g.,~K-Nearest Neighbors methods, local regression) which rely heavily on the the information content of the neighbourhood. RFI - which only partially relies on local structure of the data - suffered marginally from high missingness. SVD and MNI, which build on more general properties of the data, and Beagle - whose learning process is fundamentally different and specific for genotype imputation - were scarcely affected by missing rates.

>>>>

\subsection{Missing rate and the ``curse of dimensionality''}
\label{sec:missing_rate_curse_dimensionality}
In general, imputation methods were relatively robust to increasing missing rates. Only KNNI, and to a lesser extent RFI, showed slowly degrading performances at very high missing rates. The divergent response of the different imputation methods to missing rate became apparent only in the most challenging scenarios, when markers with up to 40-70\% missing genotypes were allowed in the data set.\\
KNNI's decrease in imputation accuracy with increasing missing rates can be interpreted in relation to the phenomenon known as the ``curse of dimensionality'' \cite{bellman_dynamic_1957,marimont_nearest_1979}: with the same amount of data, the increasing number of parameters - the ``dimensions'' of the problem - increases and complicates the identification of k neighbors which are close enough to the data point to be classified/imputed. 
On average, the side $l$ of the hypercube to include $k$ neighbors is a function of $k$, $n$ (sample size) and $p$ (number of parameters): $l = \left ( \frac{k}{n} \right )^{1/p}$. With $n$ constant, the hypercube in which the $k$ neighbors lie gets bigger as $p$ increases. This holds especially for local learning methods (e.g.,~K-Nearest Neighbors methods, local regression) which rely heavily on the the information content of the neighborhood. RFI - which only partially relies on local structure of the data - suffered marginally from high missingness. SVD and MNI, which build on more general properties of the data, and Beagle - whose learning process is fundamentally different and specific for genotype imputation - were scarcely affected by missing rates.

### 2016-02-04 inputs/materials_and_methods.tex
phrasing

\subsection{Imputation methods}
\label{sec:imputation_methods}
We considered six imputation methods: Mean Value Imputation (MNI), K-Nearest Neighbour Imputation (KNNI), Random Forest Imputation (RFI), Singular Value Decomposition Imputation (SVDI), Localized Haplotype Clustering Imputation (``Beagle'') and \textcolor{red}{haplotype reconstruction around the recombination site}
(``FILLIN'' from Tassel framework). For all algorithms we imputed a $m\times n$ matrix of m individuals and n markers whose data points, defined in the set \{0,1,2,NA\}, represented the three possible genotypes (AA, AB, and BB) and the missing value, respectively. 
\paragraph{MNI}
\label{par:MNI}
: in Mean Value Imputation each missing data point was replaced by the mean of the non-missing values for that marker, then discretized to the closest value in \{0,1,2\}.
\paragraph{KNNI}
\label{par:KNNI}
: in K-Nearest Neighbors Imputation missing data points were imputed based on the weighted average of the K closest markers \cite{troyanskaya_missing_2001} defined by the Simple Matching Coefficient distance function \cite{schwender_statistical_2007}, specifically designed for categorical data. $K=4$ was used in KNNI, and neighbour values were weighted by the reciprocal of their distance from the data point to be imputed.
\paragraph{SVDI}
\label{par:SVDI}
: Singular Value Decomposition (SVD) imputation was based on the following factorization of the genotype matrix M (m markers, n individuals):

>>>>

\subsection{Imputation methods}
\label{sec:imputation_methods}
We considered six imputation methods: Mean Value Imputation (MNI), K-Nearest Neighbors Imputation (KNNI), Random Forest Imputation (RFI), Singular Value Decomposition Imputation (SVDI), Localized Haplotype Clustering Imputation (``Beagle'') and \textcolor{red}{haplotype reconstruction around the recombination site}
(``FILLIN'' from Tassel framework). For all algorithms we imputed a $m\times n$ matrix of m individuals and n markers whose data points, defined in the set \{0,1,2,NA\}, represented the three possible genotypes (AA, AB, and BB) and the missing value, respectively. 
\paragraph{MNI}
\label{par:MNI}
: in Mean Value Imputation each missing data point was replaced by the mean of the non-missing values for that marker, then discretized to the closest value in \{0,1,2\}.
\paragraph{KNNI}
\label{par:KNNI}
: in K-Nearest Neighbors Imputation missing data points were imputed based on the weighted average of the K closest markers \cite{troyanskaya_missing_2001} defined by the Simple Matching Coefficient distance function \cite{schwender_statistical_2007}, specifically designed for categorical data. $K=4$ was used in KNNI, and neighbo
rs values were weighted by the reciprocal of their distance from the data point to be imputed.
\paragraph{SVDI}
\label{par:SVDI}
: Singular Value Decomposition (SVD) imputation was based on the following factorization of the genotype matrix M (m markers, n individuals):
